---
title: "Beer Production"
output: html_document
name: metrics_beer_production
owner: mersault
metrics:
  nb_pounds:
    title: "# of Pounds produced"
    description: "Number of pounds used in US beer productions"
dimensions:
  material_type:
    title: Type
    description: Grain/non-grain
  material:
    title: Material
    description: Malt, syrup, rice, etc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  results = "both"
) ################## maybe change material to ingredients, type as primary nomenclature
```

# Get the Data

```{r}
library(tidyverse)
library(broom)
library(ggthemes)
library(tidytuesdayR)
library(readr)
library(lubridate)
library(ggridges)
library(shiny)
library(tidymetrics)
library(shinymetrics)

theme_set(theme_light())

brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-03-31/brewing_materials.csv')
beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-03-31/beer_taxed.csv')
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-03-31/brewer_size.csv')
beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-03-31/beer_states.csv')

# Or read in with tidytuesdayR package (https://github.com/dslc-io/tidytuesdayR)
# PLEASE NOTE TO USE 2020 DATA YOU NEED TO USE tidytuesdayR version ? from GitHub

# Either ISO-8601 date or year/week works!

# Install via pak::pak("dslc-io/tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2020-03-31')


brewing_materials <- tuesdata$brewing_materials


brewing_materials <- brewing_materials  %>%
  mutate(date = ymd(paste(year, month, 1))) %>%
  filter(year <2016)
```

```{r}

head(brewing_materials)
head(beer_taxed)
head(brewer_size)
head(beer_states)

```

# Common ingredients used in US beer production

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# filter for a monthly sample

brewing_materials %>%
  filter(date == max(date)) %>%
  filter(!str_detect(material_type, "Total")) %>% 
  mutate(type = fct_reorder(type, month_current)) %>%
  ggplot(aes(type, month_current, fill = material_type)) +
  geom_col() +
  coord_flip()

# Create time-series by combining month and year into date col

  # All materials

brewing_materials %>%
  filter(!str_detect(material_type, "Total")) %>%
  mutate(type = fct_reorder(type, month_current, sum)) %>%
  ggplot(aes(date, month_current, fill =type)) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Time",
       y = "Pounds used in beer production",
       fill = "Materials") +
  theme_minimal()

# For aggregates

brewing_materials %>%
  filter(str_detect(material_type, "Total.*products")) %>%
  mutate(type = fct_reorder(type, month_current, sum)) %>%
  ggplot(aes(date, month_current, fill =type)) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Time",
       y = "Pounds used in beer production",
       fill = "Materials") +
  theme_minimal() +
  theme(legend.position = "top")

```

## TidyMetrics + ShinyMetrics

### 1. `tidymetrics::cross_by_dimensions()`

```{r echo=TRUE}

materials_dimensions_crossed <- brewing_materials %>%
  filter(!str_detect(material_type, "Total")) %>%
  cross_by_dimensions(material_type, type) %>% # Creates cross-tab with "all" totals for each dimension + grand total/ cross sectioning
  summarise(nb_pounds = sum(month_current))
  
print(materials_dimensions_crossed)
```

-   Creates additional totals for aggregations or something like that I don't know how to explain just yet lol

### 2. `tidymetrics::cross_by_periods()`

```{r}
brewing_summarised <-
  brewing_materials %>%
  rename(material = type) %>%
  filter(!str_detect(material_type, "Total")) %>%
  cross_by_dimensions(material_type, material) %>%
  cross_by_periods(c("month", "quarter", "year")) %>%
  summarise(nb_pounds = sum(month_current)) %>%
  ungroup() 
  
```

-   Similar to `cross_by_dimensions()`, but is focused on period.
-   The argument automatically detects a date/time column; no column argument necessary.
-   That said it needs an argument for the period to be aggregated — in this instance, it's `c("month", "quarter", "year")`. (p/s it can even go as far as day, heck maybe even time if it suports timestamp idk)

### Graphs

##### 3.1 Time-series graph using cross_by_period/dimension [by month]

```{r}
brewing_summarised %>%
  filter(material_type == "All", material == "All", period == "month") %>%
  ggplot(aes(date, nb_pounds)) +
  geom_line(size = 1, colour = "red") +
  scale_y_continuous(labels = scales::comma) +
  xlab("Time") +
  ylab("Materials used (lb)") +
  ggtitle("Materials used in beer production (2008 - 2015)") +
  theme_light()

```

#### 3.2 Column graph using cross_by_period/dimension [by quarter]

```{r}

brewing_summarised %>%
  filter(material_type == "All", 
         period == "quarter",
         !material == "All") %>% 
  ggplot(aes(date, nb_pounds, fill = material)) + 
  geom_col(size = 1) + 
  scale_y_continuous(labels = scales::comma) + xlab("Time") +
ylab("Materials used (lb)") + ggtitle("Materials used in beer production (2008 -
2015)") + 
  theme_light()

```

#### 3.3 Column graph using cross_by_period/dimension [by year]

```{r}
brewing_summarised %>%
  filter(material_type == "All", 
         period == "year") %>% 
  mutate(material = fct_reorder(material, nb_pounds, .desc = FALSE)) %>%
  ggplot(aes(date, nb_pounds, fill = material)) + 
  geom_col(size = 1) + 
  scale_y_continuous(labels = scales::comma) + xlab("Time") +
ylab("Materials used (lb)") + ggtitle("Materials used in beer production (2008 -
2015)") + 
  labs(fill = "Materials") +
  theme_clean() +
  theme(legend.position = "bottom")
```

### 4. `shinymetrics::use_metric_scaffold` + `shinymetrics::create_metrics`

```{r}
# use_metrics_scaffold(brewing_summarised)
brewing_metrics <- brewing_summarised %>%
  create_metrics()

```
```{r}
library(dplyr)
brewing_summarised %>% 
  group_by(material_type, material) %>% 
  summarise(nb_pounds = sum(nb_pounds), .groups = 'drop')
library(tidymetrics)
metrics_tbl <- use_metrics_scaffold(brewing_summarised, "metrics.yaml")
str(metrics_tbl)
head(metrics_tbl)
metrics_tbl <- create_metrics(metrics_tbl)
```










